Blaming the Typeless: \\ Scalable, Human-Centric Python Localization

### 2018-08-24 paper/main.tex
paper

One of the key difficulties in localizing type errors, especially for
dynamically-typed programs, is that there are often multiple
logically-valid ways to address a type error~\cite{FIXME} --- only one of
which may be desired by the programmer~\cite{FIXME}.  Consider the program
in \figref{motex}, adapted from our dataset of human submissions. The
program attempts to carry out some simple arithmetic based on a given
number and the current year. When executed, however, the program raises an
exception on line 4, related to the addition of an integer variable to a
string variable.

@@

Our raw data consist of every Python 3 program that a user executed on
PythonTutor.com~\cite{Guo2013-vu} (not in ``live'' mode) during 2017, other
than those with syntax errors or undefined variables.  Each program which
crashes (throws an uncaught Python exception) is paired with the next
program (by the same user) that does not crash, under the assumption that
the latter is the fixed version of the former. We discard pairs where the
difference between crashing and fixed versions is too high (more than a
stddev above average), since these are most likely to be violations of that
assumption (i.e., the program that does not crash is unrelated to the
crashing program). We also discard submissions that are out of scope given
PythonTutor's policies (e.g., very long-running executions or those that
use forbidden libraries).

@@

Our algorithm is based on supervised machine learning and thus requires
labeled training instances.  For these experiments, we employ a random
forest our machine learning algorithm, finding it to provide the best
accuracy on our dataset with a scalable training and testing time.  In
addition, our evaluations require a ground truth notion of which terms
correspond to correct fault localizations.

We use the terms changed involved in the fixes by actual users as our
ground truth and training information.  Many PythonTutor users use the site
in an iterative fashion: they start out by writing a program that crashes,
and then edit it until it no longer crashes. Our dataset contains only
those crashing programs for which the same user later submitted a program
that did not crash. We compute a tree-diff~\cite{tree-diff} between the
originaly, buggy submission and the first fixed submission. We define
the \emph{ground truth} correct answer to be the set of terms in the
crashing program that also appear in the diff. We discuss the implications
of this choice in \secref{threats}.

Given that notion of ground truth, a candidate fault localization answer is
\emph{accurate} if it is contained in the ground truth set. That is, if the
human user changed lines $X$ and $Y$, a technique (either our algorithm
or a baseline) is given credit for returning either $X$ or $Y$. When
considering ranked lists, a response list is \emph{top-$k$ accurate} if any one
of the top $k$ answers is contained in the ground truth set. \fixme{Read
this paragraph and correct it for truth; it is currently simply WRW's best
guess and I don't actually know what you did.}

@@

Some features, such as ``program crashes here'' and features
related to term size, are fairly direct. Others, such as ``Child3 is
statement'', a contextual, syntactic feature that was found to be very
important, merit additional explanation. In practice, we found it to
encode FIXME, since Python terms with statements in the third child node
position are FIXME. Similarly, features that determine if the parent node
is a list or dictionary are relevant because of FIXME. In beginner-written
Python we find that defects and typos often in the immediate child position
of a list or dictionary term because of FIXME. Finally, we note the
importance of our \lt{Unknown} type feature for expressions that are never
evaluated in the given trace: we find that they are relevant because of
FIXME. \fixme{Ben (etc.) must actually fill in real explanations here.}

>>>>

One of the key difficulties in localizing type errors, especially for
dynamically-typed programs, is that there are often multiple
logically-valid ways to address a type error --- only one of
which may be desired by the programmer~\cite{learning-to-blame}.  Consider the program
in \figref{motex}, adapted from our dataset of human submissions. The
program attempts to carry out some simple arithmetic based on a given
number and the current year. When executed, however, the program raises an
exception on line 4, related to the addition of an integer variable to a
string variable.

@@

Our raw data consist of every Python 3 program that a user executed on
PythonTutor.com~\cite{Guo2013-vu} (not in ``live'' mode) during 2017, other
than those with syntax errors or undefined variables.  Each program which
crashes (throws an uncaught Python exception) is paired with the next
program by the same user that does not crash, under the assumption that
the latter is the fixed version of the former. We discard pairs where the
difference between crashing and fixed versions is too high (more than a
standard deviation above average), since these are most likely to be violations of that
assumption (i.e., the program that does not crash is unrelated to the
crashing program). We also discard submissions that are out of scope given
PythonTutor's policies (e.g., very long-running executions or those that
use forbidden libraries).

@@

Our algorithm is based on supervised machine learning and thus requires
labeled training instances.  For these experiments, we employ a random
forest as our machine learning algorithm, finding it to provide the best
accuracy on our dataset with a scalable training and testing time.  In
addition, our evaluations require a ground truth notion of which terms
correspond to correct fault localizations.

We use the terms changed in fixes by actual users as our
ground truth and training information.  Many PythonTutor users use the site
in an iterative fashion: they start out by writing a program that crashes,
and then edit it until it no longer crashes. Our dataset contains only
those crashing programs for which the same user later submitted a program
that did not crash. We compute a tree-diff~\cite{tree-diff} between the
original, buggy submission and the first fixed submission. For example,
if the expression \lt{len(\{3,4\})} is changed to \lt{len([3,4])}, then the
node in the tree corresponding to the set \lt{\{3,4\}} as a whole will
appear in the diff (since it has been changed to a list), but neither its parent
node (the enclosing \lt{len} call) nor its children nodes (the integer literals)
will appear since they were not changed. Changing the number of children of a
node also causes the node to appear in the diff, e.g. adding a statement
to a while loop counts as a change to the loop as a whole.

We define the \emph{ground truth} correct answer to be the set of terms in the
crashing program that also appear in the diff. We discuss the implications
of these choices in \secref{threats}.

Given that notion of ground truth, a candidate fault localization answer is
\emph{accurate} if it is contained in the ground truth set. That is, if the
human user changed terms $X$ and $Y$, a technique (either our algorithm
or a baseline) is given credit for returning either $X$ or $Y$. When
considering ranked lists, a response list is \emph{top-$k$ accurate} if any one
of the top $k$ answers is contained in the ground truth set.

@@

Some features, such as ``program crashes here'' and features
related to term size, are fairly direct. Others, such as ``Child3 is
statement'', a contextual, syntactic feature that was found to be very
important, merit additional explanation. In practice, we found it to
encode complex control structures, like large loops, where fixing the crash
often involved adding or removing whole statements from the loop.
Similarly, features that determine if the parent node
is a list or dictionary are relevant because of FIXME. In beginner-written
Python we find that defects and typos often in the immediate child position
of a list or dictionary term because of FIXME. Finally, we note the
importance of our \lt{Unknown} type feature for expressions that are never
evaluated in the given trace: we find that they are relevant because of
FIXME. \fixme{Ben (etc.) must actually fill in real explanations here.}
